PANCHANGE_data_preprocessing

### 2020-09-01 PANCHANGE_data_preprocessing2020-07-15.Rmd
added dont know pnta for NBR ED because of screening question

edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    mhd.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa,
    mhd.atypical_bulimia_nervosa,
    mhd.bingeeating_disorder,
    mhd.atypical_bingeeating_disorder,
    mhd.ed_none_of_the_above = mhd.none_of_the_above
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.mhd.raw.id <- nbr.mhd.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd."),
    mhd.specific_phobia_e.g._phobia_of_flying = mhd.specific_phobia_e.g._fear_of_flying,
    mhd.type_psychosis_psychotic_illness = mhd.psychosis_type_psychotic_illness,
    mhd.autism_aspergers_or_autistic_spectrum_disorder = mhd.autism_spectrum_disorder_asd,
    mhd.anorexia_nervosa_unc = mhd.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa_unc = mhd.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa_unc = mhd.bulimia_nervosa,
    mhd.bingeeating_disorder_unc = mhd.bingeeating_disorder,
    mhd.ed_none_of_the_above = mhd.none_of_the_above.2
  ) %>%
  add_column(
    mhd.schizoaffective_disorder = NA #NBR doesn't ask about schizoaffective disorder. This means the column das not exist in the data set
  ) %>% 
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

There is routing questionnaire if participants have been diagnosed with an eating disorder and afterwards they get asked the specific disorders. Therefore, mhd.suspected_eating_disorder_diagnosed_numeric needs to be recoded as 0 otherwise they would be NAs.
```{r nbr cleaning eating disorders}
if(NBR == TRUE) {
#Recode as 0 if ppt responds no to screening question (mhd.suspected_eating_disorder_diagnosed_numeric)
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    mhd.anorexia_nervosa_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.anorexia_nervosa_unc_numeric == "0" ~ 0,
                mhd.anorexia_nervosa_unc_numeric == "1" ~ 1),
    mhd.atypical_anorexia_nervosa_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.atypical_anorexia_nervosa_unc_numeric == "0" ~ 0,
                mhd.atypical_anorexia_nervosa_unc_numeric == "1" ~ 1),
    mhd.bulimia_nervosa_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.bulimia_nervosa_unc_numeric == "0" ~ 0,
                mhd.bulimia_nervosa_unc_numeric == "1" ~ 1),
    mhd.bingeeating_disorder_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.bingeeating_disorder_unc_numeric == "0" ~ 0,
                mhd.bingeeating_disorder_unc_numeric == "1" ~ 1),
    mhd.anorexia_nervosa =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Anorexia nervosa",
                mhd.anorexia_nervosa_unc_numeric == "0" ~ "Not Anorexia nervosa",
                mhd.anorexia_nervosa_unc_numeric == "1" ~ "Anorexia nervosa"),
    mhd.atypical_anorexia_nervosa =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Atypical anorexia nervosa",
                mhd.atypical_anorexia_nervosa_unc_numeric == "0" ~ "Not Atypical anorexia nervosa",
                mhd.atypical_anorexia_nervosa_unc_numeric == "1" ~ "Atypical anorexia nervosa"),
    mhd.bulimia_nervosa =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Bulimia nervosa",
                mhd.bulimia_nervosa_unc_numeric == "0" ~ "Not Bulimia nervosa",
                mhd.bulimia_nervosa_unc_numeric == "1" ~ "Bulimia nervosa"),
    mhd.bingeeating_disorder =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Binge-eating disorder",
                mhd.bingeeating_disorder_unc_numeric == "0" ~ "Not Binge-eating disorder",
                mhd.bingeeating_disorder_unc_numeric == "1" ~ "Binge-eating disorder")
    )
}

@@

### Controls
These individuals have no self report diagnosis of any kind that we measure. 
```{r mhd control}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    control_numeric =
      if_else(
    depressive_disorders_numeric == "0" &
    anxiety_disorders_numeric == "0" &
    eating_disorders_numeric == "0" &
    obsessive_compulsive_disorders_numeric == "0" &
    psychotic_disorders_numeric == "0" &
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric == "0" &
    mhd.posttraumatic_stress_disorder_ptsd_numeric == "0" &
    autism_spectrum_disorder_numeric == "0" &
    mhd.attention_deficit_hyperactivity_disorder_numeric == "0" &
    mhd.personality_disorder_numeric == "0" &
      mhd.prefer_not_to_answer_numeric == "0" & # added 31. August 2020
      mhd.prefer_not_to_answer.1_numeric == "0" & # added 31. August 2020
      mhd.dont_know_numeric == "0" & # added 31. August 2020
      mhd.dont_know.1_numeric == "0" & # added 31. August 2020
      mhd.other_please_tell_us_more_numeric == "0",
    true = 1, 
    false = 0, 
    NA_real_)
  )

@@

mhd.raw.id %>%
  freq(control,
       cumul = F)
```

>>>>

edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    mhd.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa,
    mhd.atypical_bulimia_nervosa,
    mhd.bingeeating_disorder,
    mhd.atypical_bingeeating_disorder,
    mhd.none_of_the_above_eating_disorder = mhd.none_of_the_above
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.mhd.raw.id <- nbr.mhd.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd."),
    mhd.specific_phobia_e.g._phobia_of_flying = mhd.specific_phobia_e.g._fear_of_flying,
    mhd.type_psychosis_psychotic_illness = mhd.psychosis_type_psychotic_illness,
    mhd.autism_aspergers_or_autistic_spectrum_disorder = mhd.autism_spectrum_disorder_asd,
    mhd.other_please_tell_us_more = mhd.other,
    mhd.anorexia_nervosa_unc = mhd.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa_unc = mhd.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa_unc = mhd.bulimia_nervosa,
    mhd.bingeeating_disorder_unc = mhd.bingeeating_disorder,
    mhd.other_eating_disorder_unc = mhd.other_eating_disorder,
    mhd.none_of_the_above_eating_disorder_unc = mhd.none_of_the_above.2,
    mhd.dont_know_eating_disorder_unc = mhd.dont_know.2,
    mhd.prefer_not_to_answer_eating_disorder_unc = mhd.prefer_not_to_answer.2
  ) %>%
  add_column(
    mhd.schizoaffective_disorder = NA #NBR doesn't ask about schizoaffective disorder. This means the column das not exist in the data set
  ) %>% 
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

There is routing questionnaire if participants have been diagnosed with an eating disorder and afterwards they get asked the specific disorders. Therefore, mhd.suspected_eating_disorder_diagnosed_numeric needs to be recoded as 0 otherwise they would be NAs.
```{r nbr cleaning eating disorders}
#Recode as 0 if ppt responds no to screening question (mhd.suspected_eating_disorder_diagnosed_numeric)
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    mhd.anorexia_nervosa_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.anorexia_nervosa_unc_numeric == "0" ~ 0,
                mhd.anorexia_nervosa_unc_numeric == "1" ~ 1),
    mhd.atypical_anorexia_nervosa_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.atypical_anorexia_nervosa_unc_numeric == "0" ~ 0,
                mhd.atypical_anorexia_nervosa_unc_numeric == "1" ~ 1),
    mhd.bulimia_nervosa_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.bulimia_nervosa_unc_numeric == "0" ~ 0,
                mhd.bulimia_nervosa_unc_numeric == "1" ~ 1),
    mhd.bingeeating_disorder_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.bingeeating_disorder_unc_numeric == "0" ~ 0,
                mhd.bingeeating_disorder_unc_numeric == "1" ~ 1),
    mhd.none_of_the_above_eating_disorder_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.none_of_the_above_eating_disorder_unc_numeric == "0" ~ 0,
                mhd.none_of_the_above_eating_disorder_unc_numeric == "1" ~ 1),
    mhd.dont_know_eating_disorder_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.dont_know_eating_disorder_unc_numeric == "0" ~ 0,
                mhd.dont_know_eating_disorder_unc_numeric == "1" ~ 1),
    mhd.prefer_not_to_answer_eating_disorder_numeric =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.prefer_not_to_answer_eating_disorder_unc_numeric == "0" ~ 0,
                mhd.prefer_not_to_answer_eating_disorder_unc_numeric == "1" ~ 1),
    mhd.other_eating_disorder_numeric = 
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ 0,
                mhd.other_eating_disorder_unc_numeric == "0" ~ 0,
                mhd.other_eating_disorder_unc_numeric == "1" ~ 1),        
    mhd.anorexia_nervosa =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Anorexia nervosa",
                mhd.anorexia_nervosa_unc_numeric == "0" ~ "Not Anorexia nervosa",
                mhd.anorexia_nervosa_unc_numeric == "1" ~ "Anorexia nervosa"),
    mhd.atypical_anorexia_nervosa =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Atypical anorexia nervosa",
                mhd.atypical_anorexia_nervosa_unc_numeric == "0" ~ "Not Atypical anorexia nervosa",
                mhd.atypical_anorexia_nervosa_unc_numeric == "1" ~ "Atypical anorexia nervosa"),
    mhd.bulimia_nervosa =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Bulimia nervosa",
                mhd.bulimia_nervosa_unc_numeric == "0" ~ "Not Bulimia nervosa",
                mhd.bulimia_nervosa_unc_numeric == "1" ~ "Bulimia nervosa"),
    mhd.bingeeating_disorder =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Binge-eating disorder",
                mhd.bingeeating_disorder_unc_numeric == "0" ~ "Not Binge-eating disorder",
                mhd.bingeeating_disorder_unc_numeric == "1" ~ "Binge-eating disorder"),
    mhd.none_of_the_above_eating_disorder =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not None of the above",
                mhd.none_of_the_above_eating_disorder_unc == "0" ~ "Not None of the above",
                mhd.none_of_the_above_eating_disorder_unc == "1" ~ "None of the above"),
    mhd.dont_know_eating_disorder =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Don't know",
                mhd.dont_know_eating_disorder_numeric == "0" ~ "Not Don't know",
                mhd.dont_know_eating_disorder_numeric == "1" ~ "Don't know"),
    mhd.prefer_not_to_answer_eating_disorder =
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Prefer not to answer",
                mhd.prefer_not_to_answer_eating_disorder_unc == "0" ~ "Not Prefer not to answer",
                mhd.prefer_not_to_answer_eating_disorder_unc == "1" ~ "Prefer not to answer"),
    mhd.other_eating_disorder = 
      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Other eating disorder",
                mhd.other_eating_disorder_unc_numeric == "0" ~ "Not Other eating disorder",
                mhd.other_eating_disorder_unc_numeric == "1" ~ "Other eating disorder")
    )
}

@@

### Controls
These individuals have no self report diagnosis of any kind that we measure. 
```{r mhd control GLAD RAMP}
if(GLAD == TRUE | RAMP == TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    control_numeric =
      if_else(
    depressive_disorders_numeric == "0" &
    anxiety_disorders_numeric == "0" &
    eating_disorders_numeric == "0" &
    obsessive_compulsive_disorders_numeric == "0" &
    psychotic_disorders_numeric == "0" &
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric == "0" &
    mhd.posttraumatic_stress_disorder_ptsd_numeric == "0" &
    autism_spectrum_disorder_numeric == "0" &
    mhd.attention_deficit_hyperactivity_disorder_numeric == "0" &
    mhd.personality_disorder_numeric == "0" &
      mhd.prefer_not_to_answer_numeric == "0" & # added 31. August 2020
      mhd.prefer_not_to_answer.1_numeric == "0" & # added 31. August 2020
      mhd.dont_know_numeric == "0" & # added 31. August 2020
      mhd.dont_know.1_numeric == "0" & # added 31. August 2020
      mhd.other_please_tell_us_more_numeric == "0",
    true = 1, 
    false = 0, 
    NA_real_)
  )

@@

mhd.raw.id %>%
  freq(control,
       cumul = F)
}
```

```{r mhd control EDGI NBR}
if(EDGI == TRUE | NBR == TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    control_numeric =
      if_else(
    depressive_disorders_numeric == "0" &
    anxiety_disorders_numeric == "0" &
    eating_disorders_numeric == "0" &
    obsessive_compulsive_disorders_numeric == "0" &
    psychotic_disorders_numeric == "0" &
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric == "0" &
    mhd.posttraumatic_stress_disorder_ptsd_numeric == "0" &
    autism_spectrum_disorder_numeric == "0" &
    mhd.attention_deficit_hyperactivity_disorder_numeric == "0" &
    mhd.personality_disorder_numeric == "0" &
      mhd.prefer_not_to_answer_numeric == "0" & # added 31. August 2020
      mhd.prefer_not_to_answer.1_numeric == "0" & # added 31. August 2020
      mhd.dont_know_numeric == "0" & # added 31. August 2020
      mhd.dont_know.1_numeric == "0" & # added 31. August 2020
      mhd.other_please_tell_us_more_numeric == "0" &
      mhd.dont_know_eating_disorder_numeric == "0",# &
#      mhd.prefer_not_to_answer_eating_disorder_numeric == "0" &
#      mhd.other_eating_disorder_numeric == "0",
    true = 1,
    false = 0, 
    NA_real_)
  )

mhd.raw.id %>%
  freq(control,
       cumul = F)
}
```
